The Regional Effects of Marine Protected Areas

### 2019-10-10 documents/ovando-regional-effects-of-mpas.Rmd
initial submission

  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies. All of these can be interpreted as a conflict of interest when evaluating fisheries policy. C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security

@@

  Healthy marine ecosystems are critical to the well-being of the planet. Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems. Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes. Using simulations, we show that the effects of marine protected areas can be highly variable and dependent on human drivers. Building off of MPA theory, we find no statistically clear effect of a network of MPAs, which our simulation analysis suggests is entirely reasonable. Our results lay a foundation for future research on the design and monitoring of marine protected areas. 

@@

  This study would not be possible without the work provided by PISCO divers over the years. DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support. 

@@

   - \usepackage{eso-pic,graphicx,transparent}

@@

  run_name: ["v5.1"]

@@

@@

<!-- why is it going to be hard ot detect regional effects of MPAs few if any empirical studies at net regional effects. does this mean that i doesn't exist, or impossible tot est, why? three fundamental stumbling blocks -->

@@

<!-- \AddToShipoutPictureFG{ -->
<!--   \AtPageCenter{% or \AtTextCenter -->
<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
<!--     }} -->
<!--   } -->
<!-- } -->

No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources. Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978]. Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes like Yellowstone or Kruger National Parks [@iucn1976;@watson2014b]. Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to  bolster fish populations throughout the region in which they are located (what we term “regional-scale effects”) [@gaines2010].

@@

<!-- **key thing: a lot more ways for small things to happen than big things ** -->

<!-- xx need to add in discussion of other outcomes not covered here, e.g. resiliency, tourism, etc. xx -->

Much of the literature on the effects of Marine Protected Areas focuses on assessing the conservation effects within the borders of protected areas [e.g. @edgar2014]. While these within-MPA effects are vitally important for protecting rare species, biodiversity, critical habitats, and often tourism, they paint an incomplete picture of the overall population effects of MPAs. The organisms within the borders of protected areas are generally part of a broader population, the biological stock, connected through adult movement and larval dispersal. If the goal of conservationists or natural resource managers is to increase the total abundance or productivity of a resource, a broader question we should ask of MPAs is not just are there more fish inside their borders, but also how have the reserves affected abundances throughout the region in which they are located? This logic, that MPAs will have conservation benefits for most species beyond their borders, is implicit in all multilateral calls for MPA expansion.

We define the regional conservation effects of MPAs as the change in total biomass of fish (summing inside and outside of MPAs) relative to the total biomass of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis). Note that this definition of regional effects is in line with the recommendations of @ferraro2018, but differs from effects measured by for example differences in outcomes before and after MPAs.

We used a simulation model to examine the regional conservation effects of MPAs under a wide array of conditions that might reasonably be expected to affect these outcomes. The model explores combinations of MPA designs (e.g., sizes, number, and placement of MPAs), life histories (e.g., growth and mortality rates, adult movement, larval dispersal, recruitment variability and autocorrelation, and timing of density dependence), and fishing fleet dynamics (e.g., degree of fishing pressure, size selectivity of the fleet, reaction of fishing pressure to MPA placement) (See Table.S1 for a complete description of simulation variables). We use this model to explore how different drivers interact to affect the conservation and fishery outcomes of MPAs.

<!-- Numerous works have used simulation modeling to explore the effects of MPAs on outcomes such as fisheries yields and stock abundance (see @fulton2014 and references within),but these prior studies have either focused on the effect of a select set of drivers (such as dispersal rates) on outcomes, or have been highly localized assessments that incorporate multiple drivers but are tuned to one specific region. Our model captures critical bio-economic features, while still being flexible and fast enough to be generalized to a broad array of scenarios. We use this model to explore how different drivers interact to affect the conservation and fishery outcomes of MPAs. -->

@@

<!-- Across all simulations the median simulated equilibrium regional conservation effect, expressed as the percentage change in total biomass with the MPAs and without the MPAs, was `r round(100*median(eqo$mpa_effect))`%, with a minimum of less than `r round(100*min(eqo$mpa_effect))`% and a maximum of over `r round(100*max(eqo$mpa_effect))`%. The median time required for the simulated MPAs to cause a 10% increase in regional biomass was approximately 10 years. -->

@@

These simple results support the intuitive conclusion that bigger MPAs on overfished populations produce large positive conservation outcomes. However, when combined with other factors such as fleet dynamics, life history, and MPA size, the conservation effects of MPAs can vary widely in magnitude and timing, a finding corroborated by empirical evidence [@starr2015;@caselle2015;@nickols2019;@villasenor-derbez2019] (Fig.\ref{conservation-effects}-B).  MPAs protecting small proportions of a stock can produce small positive effects, large positive effects, or even small negative effects. As the area protected inside MPAs increases, positive effects become more likely, but even MPAs protecting most of a population's footprint can produce conservation effects as low as 0% and as high as 200% and above depending on the context in which they are placed. The degree of depletion (i.e. fishing pressure) has a clearer signal, with small conservation benefits of MPAs when depletion is low, and larger effects when depletion is high, but again even for severely depleted populations the conservation effects of MPAs can vary widely (although nearly always with population benefits).

@@

How can we empirically measure the regional effects of MPAs?  Intuition and bio-economic theory tells us that the regional conservation effects of MPAs are likely to be highly context dependent across locations and across species within locations, so careful thought must be devoted to an empirical design to detect their effects. The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.  In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented. Both worlds would be tracked before and after MPA implementation, and the conservation and fishery outcomes would be compared after treatment. Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.  Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are never placed at random. Careful causal inference requires some other means of controlling for biases introduced by factors such as the MPA siting process and biological spillover and concentration of fishing effort  [@ferraro2018;@larsen2019].

<!-- From the fisheries perspective, catch rates before and after [e.g. @roberts2001a] or near and far from MPAs [e.g. @russ2004a] are sometimes presented as evidence for the fisheries effects of MPAs. However neither of these approaches can be counted on to provide reliable estimate of the net fishery effects of MPAs, due to failures to account for potential broader trends in catch rates independent of MPAs and the potentially confounding effects of effort redistribution. @smith2006a presents an econometrically robust method for estimating the fishery effects of MPAs, though this approach requires high-resolution data on comparable fishing effort both near and far from MPAs, data that is very rarely available. It is beyond the scope of this paper to delve deeper into detecting the net fishery effects of MPAs, and so we now focus on the question of estimating their regional conservation effects. -->

@@

<!-- However, even under this ideal example the response ratio is not a perfect estimator due to the fact that this ratio contains no information on the effect of the MPAs outside their borders.  -->

What happens when the “control” sites are in fact affected by the application of the MPA through biological spillover and fishing concentration? In order to address this possibility, we calculated a new response ratio as the mean density inside MPAs relative to the mean density outside the MPAs (both weighted by distance from MPA borders and scaled by MPA size). We only include simulations in which habitat and larval dispersal rates are identical inside and outside of MPAs, to approximate a scenario in which treatment and control sites have been paired by ecological characteristics. Under these circumstances the response ratio is a somewhat biased and very inaccurate estimator of the true regional effect of MPAs. While on average high response ratios correspond with high true regional effects, the variance is high, so the response ratio alone cannot be used as a reliable proxy for the true regional effect, without accounting for the sources of biases discussed here. For higher movement rates, response ratios were frequently near 0\% (potentially leading stakeholders to conclude that the MPA had been ineffective), when in fact in many cases the true effect of the MPA was highly positive. For highly sedentary species, extremely high response ratios could create the appearance of massive conservation gains when in fact the net effect on the regional population has been near zero or even negative (Fig.\ref{density-ratio}-B).

@@

To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects. In reality these conditions are unlikely to hold but for a few select cases. This become especially difficult when networks of MPAs are implemented, such as has occurred in California. In this case, control sites are generally located within a few kilometers from MPA borders, making biological and economic spillover highly likely. In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas.

@@

dr_plot <- density_ratios %>%
  filter(sprinkler == FALSE, mpa_habfactor == 1) %>%
  select(experiment,
         years_protected,
         mpa_effect,
         adult_movement,
         mpa_size,
         contains("density_ratio")) %>%
  mutate(adult_movement = adult_movement / max(adult_movement)) %>%
  gather(scenario, density_ratio, contains("density_ratio")) %>%
  group_by(experiment, scenario) %>%
  filter(years_protected == max(years_protected)) %>%
  mutate(density_ratio = density_ratio * mpa_size) %>% 
  ggplot(aes(y = pmin(4, mpa_effect), x = pmin(4, density_ratio - 1))) +
    geom_hline(aes(yintercept = 0)) + 
  geom_vline(aes(xintercept = 0)) + 
  geom_abline(aes(slope = 1, intercept = 0), linetype = 2, size = 1.5, color = "black") +
  geom_point(aes(color = adult_movement, size = mpa_size), alpha = 0.5, show.legend = TRUE)  +
    geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ rev(scenario), labeller = labeller(`rev(scenario)` = c(biased_density_ratio = "A) Inside vs. Experiment",true_density_ratio = "B) Inside vs. Outside"))) +
  labs(y = "True % Increase in Biomass", x = "% Response Ratio") +
  # scale_x_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
  # scale_y_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
  scale_color_viridis(guide = wide_hgc, labels = scales::percent_format(accuracy = 1), name = "Fish Movement",
                      option = "cividis") + 
  scale_x_continuous(labels = percent) + 
    scale_y_continuous(labels = percent) + 
  theme(legend.position = "top") + 
  guides(size = FALSE)

pdf(file =  "figs/density-ratio.pdf",width = 8, height = 5, useDingbats = TRUE)
print(dr_plot)
dev.off()

@@

\begin{figure*}%[tbhp]
  \centering
  \includegraphics[width=.9\linewidth]{figs/density-ratio.pdf}
  \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis). Each point represents a simulation, and color represents the movement rate of fish in the simulation (expressed as a percentage of the patches a fish is likely to move in one time step). The black line shows the one:one fit line, the blue line is a linear fit of the relationship between observed response ratios and true regional change in biomass. Inside vs. Experiment simulates a true control, while Inside Vs. Outside uses a control that may be affected by the treatment.}
  \label{density-ratio}
\end{figure*}

@@

<!-- Our strategy for identifying the effect of the MPAs in the Channel Islands is to use biomass densities of non-targeted species as our control for broader factors affecting biomass densities of fishes throughout the Islands besides the effect of the MPAs. Simply examining densities of tarted species throughout the islands before and after MPAs might confuse changes in regional densities that would have occurred anyway without the MPAs with MPA effects (since at the regional scale we do not have a spatial control for the effects of the MPAs). We assume that over the 13 years of MPA protection covered in this study non-targeted species were relatively unaffected by the MPAs (i.e. indirect effects of MPAs on species interactions are negligible), and can can serve as an indicator for the overall changes in fish abundance throughout the Channel Islands that would have occurred without the MPAs. Raw densities calculated by PISCO provide visual support for this hypothesis, in that both the targeted and non-targeted species groups share common trends (Fig.\ref{pop-trends}, see Materials and Methods for statistical support for this assumption). -->

@@

Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs? In a manner similar to @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings]. Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish. Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014]. But, we must note that the analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).  Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species). The key assumptions of this approach are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI section 1.6.3), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities (the parallel-trends assumption). The advantage to this approach is that given the time-frame of the model (15 years), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios or a spatial BACI design.

@@

\begin{figure}%[tbhp]
  \centering
  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons).}
  \label{pop-trends}
\end{figure}

The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}). However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions. Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003), which provides support for the parallel trends assumption critical to the difference-in-difference method. Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014. Following 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).

@@

\begin{figure}%[tbhp]
  \centering
  \includegraphics[width=1\linewidth]{figs/did_plot.png}
  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs). MPAs are implemented in 2003 (red dashed line). Estimates are from a regression on log(biomass density), so estimated effects roughly correspond to percentage changes. Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands.}
  \label{did-plot}
\end{figure}

@@

<!-- Studies such as @dinerstein2019 call for 30% of the oceans to be places in protected areas. What does simulation exercise suggest that the effects of such a closure might be on overall population biomass densities? XX. Suppose we assumed that all fisheries are both open-access and overfished (a clearly worst case scenario as many fisheries are well managed). In that case, we see that... -->

<!-- <!-- zeros may be an artifact of "error in variables", which can bias estimates towards zero [@larsen2019] -->

@@

<!-- Our simulation results present a clear outcome from the perspective of fishery effects: Under the assumptions of our model, MPAs should not be expected to provide improvements in fisheries catches unless the stock is experiencing substantial overfishing (Fig.\ref{catch-msy-effects}-B). This general result has been discussed before [@hilborn2004], but our results provide the most comprehensive simulation testing of this concept to date. MPAs placed in overfished fisheries frequently produced catch gains of 0-25%, though larger and smaller effects were still possible. MPA size also presents an interesting tradeoff from a fisheries perspective. MPAs that cover less than 25% of the population range, which is true for nearly all MPAs currently on the planet, very rarely produced substantial fishery losses, but also never produced substantial fishery gains. Simulated MPAs  that covered between 25% & 50% of species ranges produced the potential for higher fishery gains, but also opened up risk of greater fishery loses. For MPAs covering more than 50% of a species range, fishery losses quickly become much more likely than fishery gains. -->

<!-- We only measure one aspect of fishery outcomes here (catches). MPAs can affect many other outcomes of importance to fisheries, include profits (e.g. if institutions do not create derby fishing and MPAs improve catch per unit effort along the borders of reserves), resiliency, and research. Our model does not account for potentially important factors such as species interactions Overall, our fishery-specific results suggest a relatively simple rule of thumb: unless there is clear evidence of overfishing, MPAs should not be expected to produce improvements in fishery catches without clear explanation for the mechanism for these increases (e.g. improvements in habitat). In assessing the potential fishery benefits of MPAs, stakeholders must make explicit the mechanisms by which these benefits are expected to be produced (for example through increased catchability of fish near MPA borders), and what evidence exists for the existence of those mechanisms within the local context. -->

We find regional conservation gains from MPAs in `r percent(pos_runs, accuracy = 1)` of our simulations (Fig.\ref{conservation-effects}). MPAs covering less than 25% of the population range were unlikely to produce regional conservation gains above 10% unless the stock would have been severely overfished without the MPA. Given that few marine species on the planet have more than 25% of their entire range protected in MPAs, this suggests that large regional gains are unlikely to be common unless the overall extent of MPAs grows globally. Large MPAs protecting highly depleted stocks almost always produced large regional conservation gains, as should be expected. The median conservation effect of `r  round(100*median(eqo$mpa_effect))`\% stands in contrast to the large within-MPA effects on biomass densities reported by @halpern2003a and @lester2009. This difference is due in part to the fact that within-MPA effects are likely to be much larger than regional-MPA effects, but may also reflect the many challenges of translating response ratios into regional conservation effects. Our simulation analysis shows that without proper controls (in particular for the effects of fleet redistribution), response ratios well over 100% are entirely plausible even when the actual conservation effect is much lower (and *vice versa*).

<!-- While our results provide important insight on the expected fishery and conservation effects of MPAs, they also highlight the critical importance of explicitly modeling the links between the biological effects of MPAs and the ways that humans respond to them. One of our clearest findings was that MPAs can cause net conservation losses when implemented in a "constant-catch" fishery, as fishers must dramatically increase their fishing effort in the waters still available to them to maintain their desired landings. The focus of much of the MPA design and modeling literature has historically been on biological and ecological such as growth rates and connectivity [citation], though studies such as @hastings1999 and @sanchirico2006b incorporate assumptions about outside-MPA fleet dynamics as well. Our results make clear that economic and ecological drivers must be included in any attempt assess the effects of MPAs MPAs: in particular, we need to explicitly understand the the behavior of fishing fleets if we hope to estimate the conservation or fishery effects of MPAs [see @costa2013 for an example of this process]. Large-scale empirical evidence confirms that predicting the effects of MPAs depends on understanding the human context in which they find themselves [@cinner2018].  -->

<!-- The question of what is a "meaningful" conservation gain is of course context dependent and depends on the priorities and preferences of stakeholders. However, under any set of preferences and objectives managers and stakeholders would of course like to know whether policy interventions such as MPAs are having their intended effects. While the importance of say a 0-10% regional conservation gain will vary, there can be little doubt that an effect size within this range will be very difficult to detect in a noisy natural system (xx citations on noise and effect size xx), and our empirical assessment of MPA conservation effects in the Channel Islands demonstrates this.  -->

<!-- . That we were unable to detect a significant divergence in the biomass densities of targeted species throughout the Channel Islands post MPA implementation, relative to the densities that the trends of the non-targeted species would suggest we should observe, should not come as a great surprise then.  -->

@@

However, our simulation analysis suggests that this result should perhaps be expected. The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished. Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains. Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{did-plot}). Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.

>>>>

  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies. All of these can be interpreted as a conflict of interest when evaluating fisheries policy. C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security. Funding for D.O. was partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.

@@

  Healthy marine ecosystems are critical to the well-being of the planet. Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems. Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes. Using simulations, we show that the regional effects of marine protected areas can be highly variable and dependent on human drivers. We find no statistically clear effect of a network of MPAs after 13 years of protection, which our simulation analysis suggests is actually to be expected. Our results lay a foundation for future research on the design and monitoring of marine protected areas. 

@@

  Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship. This study would not be possible without the work provided by PISCO divers over the years. DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support. 

@@

   - \usepackage{eso-pic,graphicx,transparent,xcolor}

@@

  run_name: ["v5.0"]

@@

linkcolor: blue

@@

@@

No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources. Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978]. Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes like Yellowstone or Kruger National Parks [@iucn1976;@watson2014b]. Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to  bolster fish populations throughout the region in which they are located [@gaines2010].

@@

Much of the literature on the effects of MPAs focuses on assessing the conservation effects within the borders of protected areas [e.g. @edgar2014]. While these within-MPA effects are vitally important for protecting rare species, biodiversity, critical habitats, and often tourism, they paint an incomplete picture of the overall population effects of MPAs. The organisms within the borders of protected areas are generally part of a broader biological stock, connected through adult movement and larval dispersal. If the goal of conservationists or natural resource managers is to increase the total abundance or productivity of a resource, a broader question we should ask of MPAs is not just are there more fish inside their borders, but also how have the reserves affected abundances throughout the region in which they are located? This logic, that MPAs will have conservation benefits for most species beyond their borders, is implicit in all multilateral calls for MPA expansion.

We define the regional conservation effects of MPAs as the change in total biomass of fish (summing inside and outside of MPAs) relative to the total biomass of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis).

We used a simulation model to examine the regional conservation effects of MPAs under a wide array of conditions that might reasonably be expected to affect these outcomes. The model explores combinations of MPA designs, life histories, and fishing fleet dynamics (See Table.S1 for a complete description of simulation variables). We use this model to explore how different drivers interact to affect the conservation outcomes of MPAs.

@@

@@

These simple results support the intuitive conclusion that bigger MPAs on overfished populations produce large positive conservation outcomes. However, when combined with other factors such as fleet dynamics, life history, and MPA size, the conservation effects of MPAs can vary widely in magnitude and timing, a finding corroborated by empirical evidence [@starr2015;@caselle2015;@nickols2019;@villasenor-derbez2019] (Fig.\ref{conservation-effects}-B). As the area protected inside MPAs increases, positive effects become more likely, but even MPAs protecting most of a population's footprint can produce conservation effects as low as 0% and as high as 200%. The degree of depletion (i.e. fishing pressure) has a clearer signal, with small conservation benefits of MPAs when depletion is low, and larger effects when depletion is high, but again even for severely depleted populations the conservation effects of MPAs can vary widely (although nearly always with population benefits).

@@

How can we empirically measure the regional effects of MPAs?  Intuition and bio-economic theory tells us that the regional conservation effects of MPAs are likely to be highly context dependent across locations and across species within locations, so careful thought must be devoted to an empirical design to detect their effects. The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.  In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented. Both worlds would be tracked before and after MPA implementation, and the conservation and fishery outcomes would be compared after treatment. Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.

Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are never placed at random. Careful causal inference requires some other means of controlling for biases introduced by factors such as the MPA siting process and biological spillover and concentration of fishing effort  [@ferraro2018;@larsen2019].

@@

What happens when the “control” sites are in fact affected by the application of the MPA through biological spillover and fishing concentration? In order to address this possibility, we calculated a new response ratio as the mean density inside MPAs relative to the mean density outside the MPAs (both weighted by distance from MPA borders and scaled by MPA size). We only include simulations in which habitat and larval dispersal rates are identical inside and outside of MPAs, to approximate a scenario in which treatment and control sites have been paired by ecological characteristics. Under these circumstances the response ratio is a somewhat biased and very inaccurate estimator of the true regional effect of MPAs.

While on average high response ratios correspond with high true regional effects, the variance is high, so the response ratio alone cannot be used as a reliable proxy for the true regional effect, without accounting for the sources of biases discussed here. For higher movement rates, response ratios were frequently near 0\% (potentially leading stakeholders to conclude that the MPA had been ineffective), when in fact in many cases the true effect of the MPA was highly positive. For highly sedentary species, extremely high response ratios could create the appearance of massive conservation gains when in fact the net effect on the regional population has been near zero or even negative (Fig.\ref{density-ratio}-B).

@@

<!-- To the extent that we can select control sites that are both sufficiently similar to the region treated with MPAs and are unaffected by biological or economic spillover, response ratios may be reliable estimators of regional conservation effects. In reality these conditions are unlikely to hold but for a few select cases. This becomes especially difficult when networks of MPAs are implemented, such as has occurred in California. In this case, control sites are generally located within a few kilometers from MPA borders, making biological spillover and fishing concentration highly likely. In these circumstances, without robust statistical controls, response ratios may provide an unreliable estimate of the regional conservation effect of marine protected areas. -->

@@

ylabs <- c(expression(""<= "-100%"), paste0(seq(0,300, by = 100),"%"),expression("">= "400%") )

dr_plot <- density_ratios %>%
  filter(sprinkler == FALSE, mpa_habfactor == 1) %>%
  select(experiment,
         years_protected,
         mpa_effect,
         adult_movement,
         mpa_size,
         contains("density_ratio")) %>%
  mutate(adult_movement = adult_movement / max(adult_movement)) %>%
  gather(scenario, density_ratio, contains("density_ratio")) %>%
  group_by(experiment, scenario) %>%
  filter(years_protected == max(years_protected)) %>%
  mutate(density_ratio = density_ratio * mpa_size) %>% 
  ggplot(aes(y = pmin(4, mpa_effect), x = pmin(4, density_ratio - 1))) +
    geom_hline(aes(yintercept = 0)) + 
  geom_vline(aes(xintercept = 0)) + 
  geom_abline(aes(slope = 1, intercept = 0), linetype = 2, size = 1.5, color = "black") +
  geom_point(aes(color = adult_movement, size = mpa_size), alpha = 0.5, show.legend = TRUE)  +
    geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ rev(scenario), labeller = labeller(`rev(scenario)` = c(biased_density_ratio = "A) Inside vs. Experiment",true_density_ratio = "B) Inside vs. Outside"))) +
  labs(y = "True % Increase in Biomass", x = "% Response Ratio") +
  # scale_x_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
  # scale_y_percent(labels = c(percent(seq(-1,1, by = 1)),">200%")) +
  scale_color_viridis(guide = wide_hgc, labels = scales::percent_format(accuracy = 1), name = "Fish Movement",
                      option = "cividis") + 
  scale_x_continuous(labels = ylabs) + 
    scale_y_continuous(labels = ylabs) + 
  theme(legend.position = "top") + 
  guides(size = FALSE)

  ggsave(dr_plot,filename = "figs/density-ratio.png",
         width = 8,
         height = 5)

# pdf(file =  "figs/density-ratio.png",width = 8, height = 5, useDingbats = TRUE)
# print(dr_plot)
# dev.off()

@@

\begin{figure*}%[tbhp]
  \centering
  \includegraphics[width=.9\linewidth]{figs/density-ratio.png}
  \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis). Each point represents a simulation, and color represents the movement rate of fish in the simulation (expressed as a percentage of the patches a fish is likely to move in one time step). The black line shows the one:one fit line, the blue line is a linear fit of the relationship between observed response ratios and true regional change in biomass. Inside vs. Experiment simulates a true control, while Inside Vs. Outside uses a control that may be affected by the treatment.}
  \label{density-ratio}
\end{figure*}

@@

@@

Given these challenges, what is a viable alternative for estimating the conservation effects of MPAs? In a manner similar to @caselle2015, we use a difference-in-difference estimator comparing densities in finfish species targeted by fishing effort (i.e., those potentially affected by an MPA) from those not targeted before and after MPA implementation [see @larsen2019 for a succinct explanation of difference-in-difference estimators in ecological settings]. Targeted species in the Northern Channel Islands, California, include commercially important finfish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish. Each of these targeted species was the subject of prior bio-economic modeling related to the effects of MPAs in southern California [@costello2010d;@rassweiler2014]. But, we must note that the analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*).  Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*) (See Table.S2 for a complete list of species). The key assumptions of this approach are that a) within the time-frame of the model there are no significant interaction effects between the targeted and non-targeted species (which in fact we do not detect, see SI), and b) that in the absence of the MPAs both the targeted and non-targeted groups of species would have exhibited similar trends in densities. The advantage to this approach is that given the time-frame of the model (16 years), we believe that spillover effects of MPA placement on the non-targeted species are likely to be much less severe than the effects of biological spillover and fishing concentration that may bias the performance of estimators such as a response ratios or a spatial BACI design.

@@

\begin{figure}%[tbhp]
  \centering
  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons). Vertical red dashed line shows year of MPA implementation.}
  \label{pop-trends}
\end{figure}

The raw trends in the biomass densities appear to show parallel trends in the densities of targeted and non-targeted fishes pre-MPA, with the targeted group diverging first positively and then negatively over time (Fig.\ref{pop-trends}). However, these raw trends could be caused by any number of confounding variables besides the effects of the MPAs themselves, from changes in observer skill and sampling locations, to shifts in environmental conditions. Using our difference-in-difference model in an effort to control for these factors, we found no significant difference in the biomass densities of targeted and non-targeted fishes pre-MPA (before 2003). Following the implementation of the MPAs in 2003, we see evidence of an increasing trend in densities of targeted relative to non-targeted fish, reaching a peak value in 2014. After 2014, densities of targeted species appear to have declined relative to what we would expect based on the non-targeted group (Fig.\ref{did-plot}).

@@

\begin{figure}%[tbhp]
  \centering
  \includegraphics[width=1\linewidth]{figs/did_plot.png}
  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs). MPAs are implemented in 2003 (red dashed line). Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
  \label{did-plot}
\end{figure}

@@

@@

We find regional conservation gains from MPAs in `r percent(pos_runs, accuracy = 1)` of our simulations (Fig.\ref{conservation-effects}). MPAs covering less than 25% of the population range were unlikely to produce regional conservation gains above 10% unless the stock would have been severely overfished without the MPA. Given that few marine species on the planet have more than 25% of their entire range protected in MPAs, this suggests that large regional gains are unlikely to be common unless the overall extent of MPAs grows globally. Large MPAs protecting highly depleted stocks almost always produced large regional conservation gains, as should be expected.

The median conservation effect of `r round(100*median(eqo$mpa_effect))`\% stands in contrast to the large within-MPA effects on biomass densities reported by @halpern2003a and @lester2009. This difference is due in part to the fact that within-MPA effects are likely to be much larger than regional-MPA effects, but may also reflect the many challenges of translating response ratios into regional conservation effects. Our simulation analysis shows that without proper controls (in particular for the effects of fleet redistribution), response ratios well over 100% are entirely plausible even when the actual conservation effect is much lower (and *vice versa*).

@@

Our simulation analysis suggests that this result should perhaps be expected. The Channel Islands MPAs cover approximately 20% of the waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished. Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains. Harkening back to our simulations, we expect the average percentage difference in densities of targeted species with and without MPAs to be modest (Fig.\ref{did-plot}). Effects of this size are likely to be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent annual survey programs such as those provided by PISCO.
